Relationships in the extremes and their influence on competition and coexistence

### 2021-09-08 paper/Paper.Rnw
editing intro

%Old version, some comments interspersed: 
%Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity, i.e., correlation coefficients.
%***DAN: "Blanketed" is a weird verb in this context. 
%***DAN: Also, MCT is a conceptually deep theory, and ATA is a fairly simple observation that happens to have some consequences. I would not be confortable with phrasing that seems to put them on the same footing! 
%Although two sets of variables may have equal correlation coefficients, they can show different association structures (see Figure).
%More specifically, the association within the tail ends of data and the asymmetry between the tail associations, termed asymmetric tail associations (ATAs), tell a more interesting story about the relationship between variables...and?...
%Recent work has shown that ATAs have important considerations for ecological phenomena; although, the implications of ATAs for competition and coexistence have not yet been investigated. 
%Asymmetry of tail associations within a dataset of variables describes uneven correlations between the tails of a dataset.
%For a simple example, a dataset with more association in the lower (upper) half and lesser association in the upper (lower) half is termed left (right) tail associated. 
%The application of ATAs in ecology is relatively new, and in exploration of ATAs in ecology, Gosh et al (AER) found ATAs in all eight ecological datasets they considered.
%Thus, the existence and functioning role of ATAs in nature is reasonable to assume and question. 
%In fact, ATAs can have influence on ecosystem service stability, Taylor’s law, and extinction risks via extreme climatic events. Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum \citep{Ghosh_2020_Eco, Ghosh_2020_AER}.
%Despite the frequent occurrence of ATAs in ecological data, the implications of tail associations for competition and coexistence, to our knowledge, has not yet been examined; doing so is the main goal of this paper.
%New version:
Every ecological discipline involves the study of relationships among biological and environmental variables,
but standard methods for examining the association between variables provide only a small portion of the overall information content contained in the
relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
the three pairs of variables in figure \ref{fig:noise} all have the same
Pearson correlation. 
%***

But each panel of figure \ref{fig:noise} nevertheless
depicts a very different type of association. Other metrics of association are used, in addition to correlations
(e.g., various common forms of regression), but correlations are extremely common, and other common
methods also provide only a small portion of the available information \citep{Nelsen2006}.
The variables in figure \ref{fig:noise}A,C 
%***
exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
they are more strongly related to each
other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}C,
termed `right-tail association') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
considered in ecology and related areas (but see \citep{Anderson_2019, deValpine_2014, Popovic_2019}), 
%***
%
our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can be mechanistically important for 
many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
%***DAN: Again note the similarity between the old and new versions, the new is just a bit more fleshed out, and maybe a
%bit more precise in some of the wordings in places.

%***DAN: Please combine the following two paragraphs into one single paragraph with this topic sentence:
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
why ATAs may be relevant to them.
%***DAN: I have inserted some comments below that might help.
Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018}].
%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects. 
In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. %**JAS: should we introduce invader analysis or terminology here?
Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones. 
For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overeall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.   
Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.

@@

Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and exent \citep{Meehl_2004, Ummenhofer_2017}. 
The rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. are a product of climate change and have implications for coexistence and competition. Changes in means can be attributed to extreme events; however changes in means are moreso a result of gradual increases or decreases over time in an environmental variable. 
Extreme events push environmental variables to their extremes, increasing variance. Effects on competition and coexistence due to climate change are more widley researched through studing changes in mean environmental changes. 
For example, decreased precipitation decreased the competitive ability of dominant forest tree species \citep{Clark_2011}. This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity as dominant species become excluded \citep{Mariotte_2013}. At the same time, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}. \cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics. Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.    
In contrast, increased environmental variance has been less commmonly studied; however, more research on the effect of increased variance has been urged and received recent focus through the study of extreme climatic events \citep{Jentsch_2007, Smith_2011}. Extreme drought and heat waves impact competitive output and community composition of grassland plants\citep{White_2000, White_2001, Hoover_2014}. However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}. Morever, in regard to invasive species, increase climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}. Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}. Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes; environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}. Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}. This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence.

Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics. If extreme events are more temporally extensive than moderate disturbances, then it is more likely that two species will be influenced by the same events. Both species may become more alike (correlated), an important consideration for competition and coexistence. For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution. This is a change in ATA not just from the event alone, but from the relationship between species. Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environment (but see Ghosh for impact of ATA on community aggregate properties).

To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple two-species lottery model. Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see figure \ref{fig:pedagog}). The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not. Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, implying less recoverablility when rare and thus less coexistence \citep{Chesson_2000} when environmentally dependent variables are asymmetrically tail associated.

To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}). The community follows the criteria required for temporal storage effects to operate. The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season. The grass dominates and has high fecundity when there is sufficeint rain in the late season. Competition is influenced by resource comsumption of growing juvenile plants. Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment. When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year. The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated. However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered. If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated. Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated. We see the possible consequences of ATA in Figure \ref{fig:pedagog}. In simulating their populations,  ATA can cause species to have longer periods of extreme dominance and lower growth rates overall.

%Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Ghosh_2020_AER}). 
%Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response \citep{Chesson_2012, Ellner_2016}.
%Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
%Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.

%If extreme events are more temporally extensive/expansive than regular environmental fluctuations, extreme events may cause the environmental response of two species who depend on temporally separated indicators of growth to become more correlated in the extremes. 
%Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics. If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events. Both species may become more alike (similar, correlated), an important consideration for competition and coexistence. For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution. This is a change in ATA not just from the event alone, but from the relationship between species. Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition, they do not directly consider ATA between species response caused by environment.

%move this paragraph to discussion
%Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
%However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers. 
%All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be. 
%Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].

This paper has the following goals. 
G1) To augment the modern theory of coexistence (MTC) to show how ATA plays a role. 
The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. 
We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
from ATAs compared to other mechanisms. This is to help illuminate whether ATA contributions to coexistence 
are likely to be negligible or meaningfully large. 
%G3) To identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model. 
%***DAN: This is a good additional goal. We will either have to have already defined the "ATA rescue and ATA exlusion sets of trait space"
%or else readers won't understand this as it was written, above (now commented). These are complex concepts, so it may be better to 
%save formal definitions for later and reword this. Here's a try at rewording: 
G3) To explore, using our simple model, whether species can coexist with
ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions to biodiversity maintenance 
may be negligible or meaninfully large.
G4) To provide an empirical example of species coexistence, and to highlight the role ATAs play in coexistence for the example.
In the Discussion, we revisit the potential that climatic changes will influence
the roles of ATAs in species coexistence and the maintenance of biodiversity.
%***DAN: I added the above sentence. Not a fully fledged "goal" but a serious thing
%to discuss.
Overall, our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climate change, 
and a means for understanding its importance for real systems.

>>>>

%Old version, some comments interspersed: 
%Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity, i.e., correlation coefficients.
%***DAN: "Blanketed" is a weird verb in this context. 
%***DAN: Also, MCT is a conceptually deep theory, and ATA is a fairly simple observation that happens to have some consequences. I would not be confortable with phrasing that seems to put them on the same footing! 
%Although two sets of variables may have equal correlation coefficients, they can show different association structures (see Figure).
%More specifically, the association within the tail ends of data and the asymmetry between the tail associations, termed asymmetric tail associations (ATAs), tell a more interesting story about the relationship between variables...and?...
%Recent work has shown that ATAs have important considerations for ecological phenomena; although, the implications of ATAs for competition and coexistence have not yet been investigated. 
%Asymmetry of tail associations within a dataset of variables describes uneven correlations between the tails of a dataset.
%For a simple example, a dataset with more association in the lower (upper) half and lesser association in the upper (lower) half is termed left (right) tail associated. 
%The application of ATAs in ecology is relatively new, and in exploration of ATAs in ecology, Gosh et al (AER) found ATAs in all eight ecological datasets they considered.
%Thus, the existence and functioning role of ATAs in nature is reasonable to assume and question. 
%In fact, ATAs can have influence on ecosystem service stability, Taylor’s law, and extinction risks via extreme climatic events. Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum \citep{Ghosh_2020_Eco, Ghosh_2020_AER}.
%Despite the frequent occurrence of ATAs in ecological data, the implications of tail associations for competition and coexistence, to our knowledge, has not yet been examined; doing so is the main goal of this paper.
%New version:
Every ecological discipline involves the study of relationships among biological and environmental variables,
but standard methods for examining the association between variables provide only a small portion of the overall information content contained in the
relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
the three pairs of variables in figure \ref{fig:noise} all have the same
Pearson correlation. 
But each panel of figure \ref{fig:noise} nevertheless
depicts a very different type of association. Other metrics of association are used, in addition to correlations
(e.g., various common forms of regression), but correlations are extremely common, and other common
methods also provide only a small portion of the available information \citep{Nelsen2006}.
The variables in figure \ref{fig:noise}A,C 
%***
exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
they are more strongly related to each
other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}C,
termed `right-tail association') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
considered in ecology and related areas (but see \citep{Anderson_2019, deValpine_2014, Popovic_2019}), 
%***
%
our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can be mechanistically important for 
many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
%***DAN: Again note the similarity between the old and new versions, the new is just a bit more fleshed out, and maybe a
%bit more precise in some of the wordings in places.

%***DAN: Please combine the following two paragraphs into one single paragraph with this topic sentence:
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
why ATAs may be relevant to them.
%***DAN: I have inserted some comments below that might help.
Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018}].
%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects. 
In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. %**JAS: should we introduce invader analysis or terminology here?
Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones. 
For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.   
Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.

@@

Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and exent \citep{Meehl_2004, Ummenhofer_2017}. 
The rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. are a product of climate change and have implications for coexistence and competition. Changes in means can be attributed to extreme events; however changes in means are moreso a result of gradual increases or decreases over time in an environmental variable. 
Extreme events push environmental variables to their extremes, increasing variance. Effects on competition and coexistence due to climate change are more widley researched through studing changes in the mean of environmental variables. 
For example, decreased growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}. This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity while dominant species become excluded \citep{Mariotte_2013}. Conversley, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}. \cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics. Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.    
In contrast, increased environmental variance has been less commmonly studied; however, more research on the effect of increased variance has been urged and received recent focus through the study of extreme climatic events \citep{Jentsch_2007, Smith_2011}. Extreme drought and heat waves impact competitive output and community composition of grassland plants\citep{White_2000, White_2001, Hoover_2014}. However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}. Morever, in regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}. Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}. Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes; environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}. Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}. This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence.

Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics. If extreme events are more temporally extensive than moderate disturbances, then it is more likely that two species will be influenced by the same events. Both species may become more alike (correlated), an important consideration for competition and coexistence. For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution. A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA). Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see Ghosh for impact of ATA on community aggregate properties).

To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple two-species lottery model, and below (in the next paragraph), we describe a hypothetical example. Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see Figure \ref{fig:pedagog}). The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not. Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, implying less recoverablility when rare and thus less coexistence \citep{Chesson_2000} when environmentally dependent variables are asymmetrically tail associated.

To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}). The community follows the criteria required for temporal storage effects to operate. The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season. The grass dominates and has high fecundity when there is sufficeint rain in the late season. Competition is influenced by resource comsumption of growing juvenile plants. Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment. When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year. The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated, and, following our simulated example (Figure \ref{fig:pedagog}), the species persist in coexistence. However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered. If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the higher values or right-tail associated. Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated. Again, we see the possible consequences of ATA in Figure \ref{fig:pedagog}, i.e., less or no coexistence.

%move this paragraph to discussion
%Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
%However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers. 
%All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be. 
%Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
We have demonstrated that ATAs exist and are common in nature \citep{Ghosh_2020_AER}, and we have elucidated how ATAs between environmental response variables might arise via extreme events. ATA in environmental response directly feeds into an ATA effect on the EC covariance contribution to the storage effect mechanism of coexistence. Thus we seek to explore the consequences of ATAs on coexistence, competition, and the maintenance of biodiversity.
This paper has the following goals. 
G1) To augment the modern theory of coexistence (MTC) to show how ATA plays a role. 
The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. 
We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
from ATAs compared to other mechanisms. This is to help illuminate whether ATA contributions to coexistence 
are likely to be negligible or meaningfully large. 
%G3) To identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model. 
%***DAN: This is a good additional goal. We will either have to have already defined the "ATA rescue and ATA exlusion sets of trait space"
%or else readers won't understand this as it was written, above (now commented). These are complex concepts, so it may be better to 
%save formal definitions for later and reword this. Here's a try at rewording: 
G3) To explore, using our simple model, whether species can coexist with
ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions to biodiversity maintenance 
may be negligible or meaninfully large.
G4) To provide an empirical example of species coexistence, and to highlight the role ATAs play in coexistence for the example.
In the Discussion, we revisit the potential that climatic changes will influence
the roles of ATAs in species coexistence and the maintenance of biodiversity.
%***DAN: I added the above sentence. Not a fully fledged "goal" but a serious thing
%to discuss.
Overall, our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climate change, 
and a means for understanding its importance for real systems. This is not a good ending sentence.
